% Define the subtitle of the page
\title{Home}

% Begin the content of the page
\subsection{A library for probabilistic modeling, inference, and criticism.}
Edward is a Python library for probabilistic modeling, inference, and
criticism. It is a testbed for fast experimentation and research with
probabilistic models, ranging from classical hierarchical models on
small data sets to complex deep probabilistic models on large data
sets. Edward fuses three fields: Bayesian statistics and machine
learning, deep learning, and probabilistic programming.


It supports \textbf{modeling languages} including
\begin{itemize}
\item \href{https://www.tensorflow.org}{TensorFlow} (with neural
  networks via \href{http://keras.io}{Keras},
  \href{https://github.com/tensorflow/tensorflow/tree/master/tensorflow/contrib/slim}{TensorFlow
  Slim}, or
  \href{https://github.com/google/prettytensor}{Pretty Tensor})
\item \href{http://mc-stan.org}{Stan}
\item \href{http://pymc-devs.github.io/pymc3/}{PyMC3}
\item Python, through \href{http://scipy.org/}{NumPy/SciPy}
\end{itemize}

It supports \textbf{inference} via
\begin{itemize}
\item Variational inference
\begin{itemize}
    \item Black box variational inference
    \item Stochastic variational inference
    \item Variational auto-encoders
    \item Inclusive KL divergence: $\text{KL}(p\|q)$
    \item Laplace approximation
  \end{itemize}
\item Marginal posterior optimization (empirical Bayes, marginal maximum
likelihood)
\item Maximum a posteriori estimation (penalized maximum likelihood, maximum
  likelihood)
\end{itemize}

It supports \textbf{criticism} of the model and inference via
\begin{itemize}
\item Point-based evaluations
\item Posterior predictive checks
\end{itemize}

Edward is built on top of
\href{https://www.tensorflow.org}{TensorFlow} as a backend, enabling
features such as computational graphs, distributed training, CPU/GPU
integration, automatic differentiation, and TensorBoard.

\subsubsection{Authors}

Edward is led by \href{http://dustintran.com}{Dustin Tran} with guidance
by \href{http://www.cs.columbia.edu/~blei/}{David Blei}. The other developers
are
\begin{itemize}
  \item \href{http://stat.columbia.edu/~diengadji/}{Adji Dieng}
  \item \href{http://www.proditus.com/}{Alp Kucukelbir}
  \item \href{http://www.ee.columbia.edu/~dliang/}{Dawen Liang}
  \item \href{http://maja-rita-rudolph.com/}{Maja Rudolph}
\end{itemize}
We welcome contributions by submitting issues, feature requests, or by solving
any current issues!

Edward has benefited enormously from the helpful feedback and advice
of many people: Jaan Altosaar, Eugene Brevdo, Allison Chaney, Matt
Hoffman, Kevin Murphy, Rajesh Ranganath, Rif Saurous, and additional
members of the Blei Lab, Google Brain, and Google Research.

\subsubsection{Citation}

We appreciate citations for Edward because it lets us find out how
people have been using the library and it motivates further work.

\begin{quote}
Dustin Tran, Adji Dieng, Alp Kucukelbir, Dawen Liang, Maja Rudolph, and David M.
Blei. 2016.
\emph{Edward: A library for probabilistic modeling, inference, and criticism. }
http://edwardlib.org
\end{quote}

\begin{lstlisting}[class=JSON]
@misc{tran2016edward,
    author = {Dustin Tran and Adji Dieng and Alp Kucukelbir and Dawen Liang and Maja Rudolph and David M. Blei},
    title = {{Edward: A library for probabilistic modeling, inference, and criticism}},
    year = {2016},
    url = {http://edwardlib.org}
}
\end{lstlisting}
