% Define the subtitle of the page
\title{Inference of probability models}

% Begin the content of the page
\subsection{Inference of probability models}

This tutorial asks the question: what does it mean to do inference in
probability models? This sets the stage for understanding how to
develop and implement inference procedures in Edward.

\subsubsection{The posterior}

How can we use a model $p(x,z)$ to analyze some data $x$? In other words,
what hidden patterns $z$ explain the data? We seek to infer these
hidden patterns using the model.

Inference for probability models leverages Bayes' rule to define the
\emph{posterior}
\begin{align*}
  p(z \mid x)
  &=
  \frac{p(x,z)}{\int p(x,z) \text{d}z}.
\end{align*}
The posterior is the distribution of the latent variables $z$, conditioned on
some (observed) data $x$. It provides a probabilistic description of the hidden
patterns that explain the data.
%Wrapping up our example, say you measured your
%weight $7$ times and the posterior of your weight turns out to be
%\begin{align*}
%  p(z \mid x) &= \mathcal{N}(z\;;\; 68.21, 2.19).
%\end{align*}
%This is a complete description of the model's estimate of your weight and its
%uncertainty around it.


\subsubsection{Inferring the posterior}

We know what the posterior represents. How do we calculate it? This is the
central computational challenge in inference.

The posterior is difficult to compute because of its normalizing
constant, which is the integral in the denominator.
This is often a high-dimensional integral that lacks an analytic (closed-form)
solution. Thus, calculating the posterior means \emph{approximating} the
posterior.

There are many approaches to posterior inference, which can be unwieldy to
manage or even conceptualize. Edward uses abstract base classes and class
inheritance to provide a hierarchy of inference methods, enabling fast
experimentation on top of existing methods. For details on the inference
hierarchy in Edward, see the
\href{api/inferences.html}{inference API}. We describe several examples in
detail in the other inference \href{tutorials.html}{tutorials}.
